Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2019-06-25 doc/network.tex
Ready to submit!!

The population-level egocentric kernel is found by integrating over individual variations:
\begin{equation}\label{eq:ego}
\hat{K}(\tau) = \int \hat{k}(\tau; a) f(a) da.
\end{equation}
\cite{trapman2016inferring} used this same kernel (also assuming a constant per-pair contact rate) to study the effect of network structure on the estimate of reproductive number.
The population-level egocentric generation-interval distribution is:
\begin{equation}
\hat{g}(\tau) = \frac{\hat{K}(\tau)}{\int \hat{K}(s) ds}.
\label{eq:conditional}
\end{equation}
The population-level egocentric generation-interval distribution describes the distribution of times at which secondary infections are realized by an \emph{average} primary case; for convenience, we will often omit ``population-level''.
Finally, the link between the growth rate and the egocentric reproductive number can be obtained by applying the Euler-Lotka equation to the egocentric generation-interval distribution \citep{trapman2016inferring}:
\begin{equation}
\frac{1}{\hat{\RR}} = \int \hat{g}(\tau) \exp(-r \tau) d\tau.
\label{eq:egorR}
\end{equation}
As the egocentric distribution always has a shorter mean than the intrinsic distribution, \Rhat\ will always be smaller than $\RR$ estimated from the intrinsic distribution;
this generation-interval-based argument provides a clear biological interpretation for the result presented by \cite{trapman2016inferring}.
% Similarly, \Rhat\ will be greater than the \emph{true} reproductive number \RR\, since it does not account for depletion of susceptibles by other routes.

For example, consider a susceptible-exposed-infected-recovered (SEIR) model, which assumes that latent and infectious periods are exponentially distributed.
The intrinsic generation-interval distribution that corresponds to this model can be written as \citep{Champredon2018equivalence}:
\begin{equation}
g(\tau) = \frac{\sigma \gamma}{\sigma - \gamma} \left(e^{-\gamma t} - e^{-\sigma t}\right),
\end{equation}
where $1/\sigma$ and $1/\gamma$ are mean latent and infectious periods, respectively.
Assuming a constant per-pair contact rate of $\lambda$ for any pair, we obtain the following egocentric generation-interval distribution:
\begin{equation}
\hat{g}(\tau) = \frac{\sigma (\gamma + \lambda)}{\sigma - (\gamma + \lambda)} \left(e^{-(\gamma + \lambda)\tau} - e^{-\sigma \tau}\right)
\end{equation}
In this case, with fixed infectiousness during the infection period, the effect of accounting for pairwise contacts is the same as an increase in the recovery rate (by the amount of the per-pair contact rate): infecting a susceptible contact is analogous to no longer being infectious (since the contact cannot be infected again); therefore, the resulting egocentric generation-interval distribution is equivalent to the intrinsic generation-interval distribution with mean latent period of $1/\sigma$ and mean infectious period of $1/(\gamma + \lambda)$.
In practice, directly using this distribution to link $r$ and $\RR$ using the Euler-Lotka equation is unrealistic because it requires that we know the per-pair contact rate. 
Instead, the per-pair contact rate can be inferred from the growth rate $r$, assuming that mean and variance of the degree distribution of a network is known (see \cite{trapman2016inferring} supplementary material section 1.4.2);
we briefly describe this relationship in Section~\ref{egosection}.
\swp{Is it OK to mention this here? Or should I say something about it at the end of the previous paragraph as well? I feel like it's too abstract to say anything there.} \jd{Seems fine this way, but let's discuss your concern.}

@@

Since the right-censoring effect is a sampling bias, we typically want to correct for it. 
In contrast, spatial effects have the same effect on how the epidemic spreads as they do on observed generation intervals. 
We therefore expect that starting from observed generation intervals and correcting for the right-censoring effect, will allow us to estimate an ``effective'' generation interval that accurately reflects dynamics of spread. 
When temporal correction is performed early in an outbreak, during the exponential growth phase, the effective distribution should incorporate egocentric and local spatial effects but not the global effects; we expect this distribution to correctly link $r$ and $\RR$. 
We will call the temporally corrected effective generation-interval distribution $\tsub{g}{eff}(\tau)$.
In a large homogeneously mixing population, the effective distribution is expected to be equivalent to the intrinsic distribution \swp{(since the egocentric depletion is negligible and the local depletion does not exist)}.
\jd{This puts a focus on how we talk about global depletion. It's sort of a thing but also sort of accounted for in the equations that contain $S$. Maybe we should adddress that point explicitly.}

>>>>

The population-level egocentric kernel is found by integrating over individual variations:
\begin{equation}\label{eq:ego}
\hat{K}(\tau) = \int \hat{k}(\tau; a) f(a) da.
\end{equation}
\cite{trapman2016inferring} used this same kernel (also assuming a constant per-pair contact rate) to study the effect of network structure on the estimate of reproductive number.
The population-level egocentric generation-interval distribution is:
\begin{equation}
\hat{g}(\tau) = \frac{\hat{K}(\tau)}{\int \hat{K}(s) ds}.
\label{eq:conditional}
\end{equation}
The population-level egocentric generation-interval distribution describes the distribution of times at which secondary infections are realized by an \emph{average} primary case; for convenience, we will often omit ``population-level''.
Finally, the link between the growth rate and the egocentric reproductive number are linked by the egocentric generation-interval distribution (and the Euler-Lotka equation) \citep{trapman2016inferring}:
\begin{equation}
\frac{1}{\hat{\RR}} = \int \hat{g}(\tau) \exp(-r \tau) d\tau.
\label{eq:egorR}
\end{equation}
As the egocentric distribution always has a shorter mean than the intrinsic distribution, \Rhat\ will always be smaller than $\RR$ estimated from the intrinsic distribution;
this generation-interval-based argument provides a clear biological interpretation for the result presented by \cite{trapman2016inferring}.
% Similarly, \Rhat\ will be greater than the \emph{true} reproductive number \RR\, since it does not account for depletion of susceptibles by other routes.

For example, consider a susceptible-exposed-infected-recovered (SEIR) model, which assumes that latent and infectious periods are exponentially distributed.
The intrinsic generation-interval distribution that corresponds to this model can be written as \citep{Champredon2018equivalence}:
\begin{equation}
g(\tau) = \frac{\sigma \gamma}{\sigma - \gamma} \left(e^{-\gamma t} - e^{-\sigma t}\right),
\end{equation}
where $1/\sigma$ and $1/\gamma$ are mean latent and infectious periods, respectively.
Assuming a constant per-pair contact rate of $\lambda$ for any pair, we obtain the following egocentric generation-interval distribution:
\begin{equation}
\hat{g}(\tau) = \frac{\sigma (\gamma + \lambda)}{\sigma - (\gamma + \lambda)} \left(e^{-(\gamma + \lambda)\tau} - e^{-\sigma \tau}\right)
\end{equation}
In this case, with fixed infectiousness during the infection period, the effect of accounting for pairwise contacts is the same as an increase in the recovery rate (by the amount of the per-pair contact rate): infecting a susceptible contact is analogous to no longer being infectious (since the contact cannot be infected again); therefore, the resulting egocentric generation-interval distribution is equivalent to the intrinsic generation-interval distribution with mean latent period of $1/\sigma$ and mean infectious period of $1/(\gamma + \lambda)$.
In practice, directly using this distribution to link $r$ and $\RR$ using the Euler-Lotka equation is unrealistic because it requires that we know the per-pair contact rate. 
Instead, the per-pair contact rate can be inferred from the growth rate $r$, assuming that mean and variance of the degree distribution of a network is known (see \cite{trapman2016inferring} supplementary material section 1.4.2);
we briefly describe this relationship in Section~\ref{egosection}.

@@

Since the right-censoring effect is a sampling bias, we typically want to correct for it. 
In contrast, spatial effects have the same effect on how the epidemic spreads as they do on observed generation intervals. 
We therefore expect that starting from observed generation intervals and correcting for the right-censoring effect, will allow us to estimate an ``effective'' generation interval that accurately reflects dynamics of spread. 
When temporal correction is performed early in an outbreak, during the exponential growth phase, the effective distribution should incorporate egocentric and local spatial effects but not the global effects; we expect this distribution to correctly link $r$ and $\RR$. 
We will call the temporally corrected effective generation-interval distribution $\tsub{g}{eff}(\tau)$.
In a large homogeneously mixing population, the effective distribution is expected to be equivalent to the intrinsic distribution (since the egocentric and local depletion are negligible in the absence of network structure, and global depletion is negligible early in the epidemic).
